CdGF_aucrf

### 2016-09-21 code/CdGF_AUCRF.Rmd
remove invsimp,cfu,entero from AUCRF

Predict_early_euth_df <- merge(meta_file[meta_file$day == 0, c('mouse_id', 'Euth_Early')], 
                               rel_abund_stool, by.x = 'mouse_id', by.y='row.names')
Predict_early_euth_df_names <- Predict_early_euth_df$mouse_id
Predict_early_euth_df <- select(Predict_early_euth_df, -mouse_id)

@@

#### 10-fold Cross validated - 10 repetitions
start <- proc.time()[3]
iters <- 100
cv10f_aucs <- c()
cv10f_all_resp <- c()
cv10f_all_pred <- c()
for(j in 1:iters){
  sampling <- sample(1:nrow(aucrf_data),nrow(aucrf_data),replace=F)
  cv10f_probs <- rep(NA,nrow(aucrf_data))
  for(i in seq(1,nrow(aucrf_data),5)){
    train <- aucrf_data[sampling[-(i:(i+4))],]
    test <- aucrf_data[sampling[i:(i+4)],]
    temp_model <- AUCRF(Euth_Early~., data=train, pdel=0.99, ntree=500)
    cv10f_probs[sampling[i:(i+4)]] <- predict(temp_model$RFopt, test, type='prob')[,2]
  }
  cv10f_roc <- roc(aucrf_data$Euth_Early~cv10f_probs)
  cv10f_all_pred <- c(cv10f_all_pred, cv10f_probs)
  cv10f_all_resp <- c(cv10f_all_resp, aucrf_data$Euth_Early)
  cv10f_aucs[j] <- cv10f_roc$auc
}
cv10f_roc <- roc(cv10f_all_resp~cv10f_all_pred)
end <- proc.time()[3]
time <- end-start

@@

#### leave-one-cage-out
LOcO_df <- merge(meta_file[meta_file$day == 0, c('mouse_id', 'cage_id', 'Euth_Early')], 
                               rel_abund_stool[ , rf_early_aucrf$Xopt], by.x = 'mouse_id', by.y='row.names')
cage_list <- unique(LOcO_df$cage_id)

LOcO_probs <- NULL
LOcO_probs <- sapply(1:length(cage_list), function(i){
  by_cage <- LOcO_df$cage_id == cage_list[i]
  LOcO_predict <- select(LOcO_df, -mouse_id, -cage_id)
  train <- LOcO_predict[-by_cage,]
  test <- LOcO_predict[by_cage,]
  temp_model <- AUCRF(Euth_Early~., data=train, pdel=0.99, ntree=500)
  LOcO_probs[i] <- predict(temp_model$RFopt, test, type='prob')[,2]
})
LOcO_probs <- unlist(LOcO_probs)
LOcO_roc <- roc(aucrf_data$Euth_Early~LOcO_probs)

@@

set.seed(seed)

@@

```{r fig.width=10, fig.height=8}
layout(matrix(c(1,1,4,4,
                2,2,3,3,
                2,2,3,3), nrow=3, byrow = TRUE))

@@

@@

#### leave-one-cage-out
LOcO_df <- cbind(cage_id = sample(meta_file[meta_file$day %in% 0, 'cage_id']),
                 cbind(Euth_Early, rel_abund_stool[ , rf_early_aucrf$Xopt]))
cage_list <- unique(LOcO_df$cage_id)

#### plot rocs
par(mar=c(4,4,1,1))
plot(c(1,0),c(0,1), type='l', lty=3, xlim=c(1.01,0), ylim=c(-0.01,1.01), xaxs='i', yaxs='i', ylab='', xlab='')
plot(otu_euth_roc, col='black', lwd=2, add=T, lty=1)
plot(cv10f_roc, col='gray', lwd=2, add=T, lty=1)
plot(LOO_roc, col='blue', lwd=2, add=T, lty=2)
plot(LOcO_roc, col='red', lwd=2, add=T, lty=2)
mtext(side=2, text="Sensitivity", line=2.5)
mtext(side=1, text="Specificity", line=2.5)
legend('bottomright', legend=c(sprintf('Out-Of-Bag (AUC: %.3g)',otu_euth_roc$auc),
                               sprintf('Leave-1 mouse-out (AUC: %.3g)',LOO_roc$auc),
                               sprintf('Leave-1 cage-out (AUC: %.3g)',LOcO_roc$auc),
                               sprintf('10-fold CV (AUC: %.3g)',cv10f_roc$auc),
                               sprintf('OOB vs Leave-1-out: p=%.2g', roc.test(otu_euth_roc,LOO_roc)$p.value),
                               sprintf('OOB vs 10-fold CV: p=%.2g', roc.test(otu_euth_roc,cv10f_roc)$p.value)
),lty=c(1,2,1,2,0,0), lwd=2, col=c('black','blue','gray','red'), bty='n')
```

```{r Random_valid OTUs, fig.width=10, fig.height=8}
layout(matrix(c(1,1,4,4,
                2,2,3,3,
                2,2,3,3), nrow=3, byrow = TRUE))

>>>>

Predict_df <- merge(meta_file[meta_file$day %in% 0, 
                              c('cage_id', 'mouse_id', 'Euth_Early')],
                    rel_abund_stool, by.x = 'mouse_id', by.y = 'row.names')

Predict_early_euth_df <- select(Predict_df, -cage_id, -mouse_id)

@@

#### 10-fold Cross validated - 10 repetitions
iters <- 100
cv10f_aucs <- c()
cv10f_all_resp <- c()
cv10f_all_pred <- c()
for(j in 1:iters){
  sampling <- sample(1:nrow(aucrf_data),nrow(aucrf_data),replace=F)
  cv10f_probs <- rep(NA,nrow(aucrf_data))
  for(i in seq(1,nrow(aucrf_data),5)){
    train <- aucrf_data[sampling[-(i:(i+4))],]
    test <- aucrf_data[sampling[i:(i+4)],]
    temp_model <- AUCRF(Euth_Early~., data=train, pdel=0.99, ntree=500)
    cv10f_probs[sampling[i:(i+4)]] <- predict(temp_model$RFopt, test, type='prob')[,2]
  }
  cv10f_roc <- roc(aucrf_data$Euth_Early~cv10f_probs)
  cv10f_all_pred <- c(cv10f_all_pred, cv10f_probs)
  cv10f_all_resp <- c(cv10f_all_resp, aucrf_data$Euth_Early)
  cv10f_aucs[j] <- cv10f_roc$auc
}
cv10f_roc <- roc(cv10f_all_resp~cv10f_all_pred)

@@

#### leave-one-cage-out
LOcO_df <- Predict_df %>% 
                select(Euth_Early, cage_id, one_of(rf_early_aucrf$Xopt))
cage_list <- as.character(unique(LOcO_df$cage_id))

@@

@@

```{r fig.width=10, fig.height=8}
layout(matrix(c(1,1,1,1,4,4,4,
                2,2,2,5,3,3,3,
                2,2,2,5,3,3,3), nrow=3, byrow = TRUE))

@@

Would adding initial increase/slope of CFU or day 1 CFU help predict?

* Adding CFU decreases AUC and improves Random model, likely due to samples lost due to lack of cfu data which are biased towards the severe samples
    - Day 1 CFU loses 2 samples  [AUCRF model with Day 1 CFU](~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/CdGF_AUCRF_withCFU_day1.html)
    - Day 2 cfu loses 14 samples, 10 are severe  [AUCRF model with Day 2 CFU](~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/CdGF_AUCRF_withCFU_day2.html)

Would it help to add community information such as diversity or enterotype?

* When adding diversity and CFU, invsimpson and cfu appear as features, but when cfu is removed, invsimpson is no longer a predictor

But how to show since model is already approaching 1?

* Rerun AUCRF with added features and see if they replace any OTUs

@@

#### leave-one-cage-out
LOcO_df <- cbind(cage_id = sample(meta_file[meta_file$day %in% 0, 'cage_id'],
                                  nrow(Validate_early_euth_df)),
                 Validate_early_euth_df[ , c('Euth_Early', rf_early_aucrf$Xopt)])
cage_list <- as.character(unique(LOcO_df$cage_id))

LOcO_probs <- NULL
LOcO_probs <- sapply(1:length(cage_list), function(i){
  by_cage <- LOcO_df$cage_id == cage_list[i]
  LOcO_predict <- select(LOcO_df, -cage_id)
  train <- LOcO_predict[-by_cage,]
  test <- LOcO_predict[by_cage,]
  temp_model <- AUCRF(Euth_Early~., data=train, pdel=0.99, ntree=500)
  LOcO_probs[[i]] <- predict(temp_model$RFopt, test, type='prob')[,2]
})
LOcO_probs <- unlist(LOcO_probs)
LOcO_roc <- roc(aucrf_data$Euth_Early~LOcO_probs)

#### plot rocs
par(mar=c(4,4,1,1))
plot(c(1,0),c(0,1), type='l', lty=3, xlim=c(1.01,0), ylim=c(-0.01,1.01), xaxs='i', yaxs='i', ylab='', xlab='')
plot(otu_euth_roc, col='black', lwd=2, add=T, lty=1)
plot(cv10f_roc, col='gray', lwd=2, add=T, lty=1)
plot(LOO_roc, col='blue', lwd=2, add=T, lty=2)
plot(LOcO_roc, col='red', lwd=2, add=T, lty=2)
mtext(side=2, text="Sensitivity", line=2.5)
mtext(side=1, text="Specificity", line=2.5)
legend('bottomright', legend=c(sprintf('Out-Of-Bag (AUC: %.3g)',otu_euth_roc$auc),
                               sprintf('Leave-1 mouse-out (AUC: %.3g)',LOO_roc$auc),
                               sprintf('Leave-1 cage-out (AUC: %.3g)',LOcO_roc$auc),
                               sprintf('10-fold CV (AUC: %.3g)',cv10f_roc$auc),
                               sprintf('OOB vs Leave-1-out: p=%.2g', roc.test(otu_euth_roc,LOO_roc)$p.value),
                               sprintf('OOB vs 10-fold CV: p=%.2g', roc.test(otu_euth_roc,cv10f_roc)$p.value)
),lty=c(1,2,2,1,0,0), lwd=2, col=c('black','blue','red','gray'), bty='n')
```

```{r Random_valid OTUs, fig.width=10, fig.height=8}
layout(matrix(c(1,1,1,1,4,4,4,
                2,2,2,5,3,3,3,
                2,2,2,5,3,3,3), nrow=3, byrow = TRUE))
